Accelerating Traffic Simulation and Its Real-Time Challenges

### 2012-09-11 draft.tex
added GPU Implementation

Deployment of traffic simulation may require real-time feedback from the
real world.
This turns out to be a cyber-physical systems (CPS) problem.
Real-time traffic simulation is another challenging issue, where
the rate and the preciseness of simulation must be traded to meet the
requirement of a given scenario.
For instance, evacuation may want very high rate simulation, sacrificing
the preciseness of simulation to some extent.
Unfortunately, none of those challenges has been explored yet, largely
due to a lack of multidisplinary research collaborations between physics
and computer science.

@@

\begin{enumerate}
 \item Initalize the time $t$, and set the initial values of $x_n(t)$
       (also $y_n(t)$ and $z_n(t)$, if necessary for multidimensional
       versions).
 \item Increase the time $t$ by the sampling period $\Delta t$.
 \item Compute the location $x_n(t)$ (also $y_n(t)$ and $z_n(t)$, if
       necessary for multidimensional versions), and the velocity
       $v_x(t)$ at time $t$ for each agent $A_n$, using the OV model.
 \item Go back to Step 2, if the simulation time is expired.
 \item Exit the program.
\end{enumerate}

Note that Step 3 explots a lot of loop procedures to derive the
locations and velocities of agents.
This part can be accelerated by the GPU.

@@

GPU performance is dominated by the program design.
GPU-accelerated programs are typically divided into two pieces of code.
The CPU code plays a role of a master thread that controls the program
flow.
The GPU code, on the other hand, spawns a bunch of worker threads to
execute compute-intensive parts of the program in parallel, thus
accelerating the overall program.
What is often argued to optimize performance is how to parallelize the
program into threads.

In order to optimize GPU performance, however, we need to make a
difficult decision about when to offload on to the GPU.

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-cpu.eps}
\caption{Sampling by the CPU.}
\label{fig:sample-in-cpu}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-gpu.eps}
\caption{Sampling by the GPU.}
\label{fig:sample-in-cpu}
\end{figure}

@@

@@

@@

Gdev~\cite{Kato2012}.

>>>>

Deployment of traffic simulation may require real-time feedback from the
real world.
This turns out to be a cyber-physical systems (CPS) problem.
Real-time traffic simulation is another challenging issue, where
the rate and the preciseness of simulation must be traded to meet the
requirement of a given scenario.
For instance, evacuation may want very high rate simulation, sacrificing
the preciseness of simulation to some extent.
Unfortunately, none of those challenges has been explored yet, largely
due to a lack of multidisciplinary research collaborations between physics
and computer science.

@@

grid, block, thread.

A unit of functions that is launched on the GPU is called a \textit{kernel}.

\section{Traffic Simulation}
\label{sec:traffic_simulation}

\begin{enumerate}
 \item Initialize the time $t$, and set the initial values of $x_n(t)$
       (also $y_n(t)$ and $z_n(t)$, if necessary for multidimensional
       versions).
 \item Increase the time $t$ by the sampling period $\Delta t$.
 \item Compute the location $x_n(t)$ (also $y_n(t)$ and $z_n(t)$, if
       necessary for multidimensional versions), and the velocity
       $v_x(t)$ at time $t$ for each agent $A_n$, using the OV model.
 \item Go back to Step 2, if the simulation time is expired.
 \item Exit the program.
\end{enumerate}

Note that Step 3 exploits a lot of loop procedures to derive the
locations and velocities of agents.
This part can be accelerated by the GPU.

@@

GPU performance is dominated by the program design.
GPU-accelerated programs are typically divided into two pieces of code.
The CPU code plays a role of a master thread that controls the program
flow.
The GPU code, on the other hand, spawns a bunch of worker threads to
execute compute-intensive parts of the program in parallel, thus
accelerating the overall program.
What is often argued in performance optimization is how to parallelize
the compute-intensive parts into threads.
This is actually the well-studied problem in the literature.
What is not really understood yet is when to offload the program on to
the GPU.
This paper explores two schemes that use the GPU at different timings to
see how GPU performance is affected.

\subsection{Sample-in-CPU Scheme}
\label{sec:sample-in-cpu}

We first implement such a scheme that uses the GPU \textit{only if
necessary}.
In other words, the program is offloaded on to the GPU, only when
computations of locations and velocities of agents are accelerated by
parallelization.
The CPU bridges across sampling periods to manage simulation.
In this scheme, the control flow of simulation is always returned to the
CPU at the end of each sampling period.
Such a synchronized approach makes the programmer easy to
obtain intermediate results of simulation, whereas the overhead imposed
on moving back and forth between the CPU and the GPU must be
compromised.

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-cpu.eps}
\caption{Block diagram of the Sample-in-CPU scheme.}
\label{fig:sample-in-cpu}
\end{figure}

Figure~\ref{fig:sample-in-cpu} shows a brief overview of the
Sample-in-CPU scheme.
This scheme also has several alternatives depending on how many kernels
need to be launched on the GPU in each period. 
Suppose that we want to offload $n$ pieces of \textit{for} loops on to
the GPU.
We may return to the CPU $n$ times in totall, that is, return every time
one \textit{for} loop breaks, or otherwise we may just return once when
all the $n$ \textit{for} loops end.
This is a design decision, and is also dependent on the GPU architecture
and the parallelization structure.
In general, parallelized threads need to be synchronized when moving
across basic blocks.
Specifically, when moving to the next \textit{for} loop executed in
parallel, all the threads relevant to this parallel computing procedure
must synchronize with each other.
However, the maximum number of threads that can be synchronized on the
GPU is often limited.
As of 2012, for example, NVIDIA's GPU architectures limit the number of
such threads to 1024 or less.
Our implementation therefore forces the program to return to the CPU
every time one \textit{for} loop breaks.
Note that this is not a conceptual limitation of GPU computing, but is a
current limitation of hardware.
We believe that this limitation would be removed or mitigated in
next-generation GPU architectures.

\subsection{Sample-in-GPU Scheme}
\label{sec:sample-in-gpu}

\begin{figure}[t]
\centering
\includegraphics[width=0.4\textwidth]{eps/sample-in-gpu.eps}
\caption{Block diagram of the Sample-in-GPU scheme.}
\label{fig:sample-in-gpu}
\end{figure}

We next implement such a scheme that uses the GPU \textit{all the time},
even to control simulation.
There is one big kernel running on the GPU, which is launched only once
at the beginning.
After offloading the program on to the GPU, the CPU is going to wait for
the completion of simulation.
This scheme is almost optimized in performance, since there is little
overhead in communication between the CPU and the GPU.
The downside of this scheme is that the CPU and the GPU are not
synchronized.
The progress of simulation is not visible from the CPU, unless the
program implements a specific interface to allow the CPU to access
intermediate results of simulation running on the GPU.
In our implementation, we consider providing a framework that the CPU
downloads data from the GPU asynchronously without awareness of
simulation, when the user requests intermediate results.
This asynchronous data access may affect the performance of simulation,
though in reality this access would not happen more than once in a
sampling period.

Figure~\ref{fig:sample-in-gpu} shows a block diagram of the
Sample-in-GPU scheme.
The procedure is very simple.
The most portion of code of simulation is executed on the GPU.
Given that the single-thread performance of recent GPUs is getting more
reliable, this approach is pretty reasonable.
As mentioned in Section~\ref{sec:sample-in-cpu}, however, the current
limitation of the GPU architecture prevents us from synchronizing among
blocks at a scale of thousands threads.
Therefore, this scheme is speculative in a sense that it does not work today
but may appear in the future.

This paper provides some degree of insights into how this scheme is
effective.
We implement this scheme with the current GPU architecture under the
assumption that global synchronization among blocks works.
There is also another possible approach for this scheme that we limit
the number of threads to what is supported by the GPU architecture.
This alternative implementation, however, is left open for future work.

@@

We evaluate the performance of our GPU-accelerated traffic simulation
programs, using an NVIDIA GeForce GTX 560 Ti graphics card.
This is a middle-end graphics card integrating 394 compute cores on a
chip.

@@

\subsection{Experimental Results}
\label{sec:experimental_results}

\subsection{Discussion}
\label{sec:discussion}

Imprecise Computation.
Gdev~\cite{Kato2012}.

@@
